---
title: "p8105_hw6_tc3326"
author: "KK Chen"
date: "2024-12-02"
output: github_document
---

```{r}
library(tidyverse)
library(modelr)
library(broom)
```

## Problem 1

download data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
# Use 5000 bootstrap samples
bootstraps <- 
  weather_df %>%
  modelr::bootstrap(5000) %>%
  mutate(
    bootstrap_data = map(strap, as_tibble),
    fitted_models = map(bootstrap_data, ~ lm(tmax ~ tmin, data = .x))
  )

# Compute estimates for R-squared, β0, and β1
weather_results <- 
  bootstraps %>%
  mutate(coefficients = map(fitted_models, broom::tidy)) %>%
  unnest(coefficients) %>%
  select(bootstrap_id = .id, fitted_models, term, estimate) %>%
  pivot_wider(
    id_cols = c(bootstrap_id, fitted_models),
    names_from = term,
    values_from = estimate
  ) %>%
  select(bootstrap_id, 
         fitted_models, 
         intercept = `(Intercept)`, 
         slope = tmin) %>%
  mutate(
    log_intercept_slope = log(intercept * slope)
  ) %>%
  mutate(model_summary = map(fitted_models, broom::glance)) %>%
  unnest(model_summary) %>%
  select(bootstrap_id, intercept, slope, log_intercept_slope, r_squared = r.squared)

# Plot for R-squared
plot_r_squared <- weather_results %>%
  ggplot(aes(x = r_squared)) +
  geom_histogram(binwidth = 0.005, fill = "blue", color = "black", alpha = 0.8) +
  labs(
    title = "Bootstrap Distribution of R-squared",
    x = "R-squared",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

plot_r_squared

# Plot for log(β0 * β1)
plot_log_beta <- weather_results %>%
  ggplot(aes(x = log_intercept_slope)) +
  geom_histogram(binwidth = 0.005, fill = "green", color = "black", alpha = 0.8) +
  labs(
    title = "Bootstrap Distribution of log(β0 * β1)",
    x = "log(β0 * β1)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

plot_log_beta
```
* The R-squared values are concentrated around 0.91, with most values between 0.89 and 0.93. The log(β₀ * β₁) centered around 2.01, with most values between 1.96 and 2.05. Both distributions are symmetric and stable, indicating the robustness of the linear regression model.

95% confidence interval for R-squared and log(β0 * β1)
```{r}
ci_r_squared <- quantile(weather_results$r_squared, probs = c(0.025, 0.975))
ci_log_intercept_slope <- quantile(weather_results$log_intercept_slope, probs = c(0.025, 0.975))

tibble(
  Metric = c("R-squared", "log(β0 * β1)"),
  `2.5%` = c(ci_r_squared[1], ci_log_intercept_slope[1]),
  `97.5%` = c(ci_r_squared[2], ci_log_intercept_slope[2])
)
```

## Problem 2 
Import and Read the data
```{r}
homicides = read_csv("./data/homicide-data.csv", na = c(".", "", "NA"))
```

Data processing
```{r}
homicides_clean <- homicides %>%
  mutate(
    city_state = paste(city, state, sep = ", "),
    solved_binary = ifelse(disposition %in% c("Closed without arrest", "Open/No arrest"), 0, 1)
  ) %>%
  filter(
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")),
    victim_race %in% c("White", "Black")
  ) %>%
  mutate(victim_age = as.numeric(victim_age)) %>%
  drop_na(victim_age)
```

Logistic Regression for Baltimore
```{r}
baltimore_model <- homicides_clean %>%
  filter(city_state == "Baltimore, MD") %>%
  glm(solved_binary ~ victim_age + victim_sex + victim_race, data = ., family = binomial)

baltimore_results <- broom::tidy(baltimore_model) %>%
  mutate(odds_ratio = exp(estimate))

baltimore_ci <- exp(confint(baltimore_model))
baltimore_or <- baltimore_results %>%
  filter(term == "victim_sexMale") %>%
  summarize(
    odds_ratio = odds_ratio,
    ci_lower = baltimore_ci["victim_sexMale", 1],
    ci_upper = baltimore_ci["victim_sexMale", 2]
  )

baltimore_or
```

Logistic Regression for All Cities
```{r}
city_model_results <- homicides_clean %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    city_model = map(data, ~ glm(solved_binary ~ victim_age + victim_sex + victim_race, data = ., family = binomial)),
    city_or_ci = map(city_model, function(model) {
      model_tidy <- broom::tidy(model) %>%
        mutate(odds_ratio = exp(estimate))
      ci <- exp(confint(model))
      sex_ci <- ci["victim_sexMale", ]
      model_tidy %>%
        filter(term == "victim_sexMale") %>%
        summarize(
          odds_ratio = odds_ratio,
          ci_lower = sex_ci[1],
          ci_upper = sex_ci[2]
        )
    })
  ) %>%
  select(city_state, city_or_ci) %>%
  unnest(city_or_ci)

city_model_results %>%
  knitr::kable(
    digits = 3,
    format = "markdown"
  )
```

Plot of ORs and CIs by City
```{r}
city_model_results %>%
  ggplot(aes(x = reorder(city_state, odds_ratio), y = odds_ratio)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
  coord_flip() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
  labs(
    title = "Estimated Odds Ratios and Confidence Intervals by City",
    caption = "Adjusted Odds Ratios for Solving Homicides (Male vs Female Victims)",
    x = "City",
    y = "Adjusted Odds Ratio (95% CI)"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, size = 14)
  )
```

* In six cities: Albuquerque, Stockton, Fresno, Nashville, Richmond, and 
Atlanta, male victims are more likely to have cases solved. In most cities, male victims are less likely to have their cases solved compared to female victims. 

## Problem 3
Import and Clean data
```{r}
birthweight_data <- read_csv("./data/birthweight.csv", na = c(".", "", "NA")) %>%
  mutate(
    baby_sex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    father_race = factor(frace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    mother_race = factor(mrace, levels = c(1, 2, 3, 4, 8), labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    malformations = factor(malform, levels = c(0, 1), labels = c("Absent", "Present"))
  )
```

Stepwise regression to propose a model
```{r}
full_model <- lm(bwt ~ ., data = birthweight_data)
stepwise_model <- step(full_model, direction = "both")

summary (stepwise_model)

final_model <- lm(bwt ~ baby_sex + bhead + blength + delwt + fincome + gaweeks + mheight + mother_race + parity + ppwt + smoken, data = birthweight_data)

summary(final_model)
```

* Started with a full model including all predictors. Used stepwise regression (forward and backward selection) to identify significant predictors. Final model included variables related to baby characteristics, maternal factors, gestational age, socioeconomic status, and parity. Achieved an adjusted R-squared of 0.7173 and residual standard error of 272.3 grams.

Add predictions and residuals to the data for plotting
```{r}
birthweight_data <- birthweight_data %>%
  add_predictions(final_model, var = "fitted_values") %>%
  add_residuals(final_model, var = "residuals")

# Plot residuals vs fitted values
ggplot(birthweight_data, aes(x = fitted_values, y = residuals)) +
  geom_point(size = 0.7, alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

Compare model to two others
```{r}
# Cross-validation setup
cv_data <- crossv_mc(birthweight_data, 100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

# Apply models to cross-validation data
cv_results <- cv_data %>%
  mutate(
    final_model = map(train, ~ lm(bwt ~ baby_sex + bhead + blength + delwt + fincome + gaweeks + mheight + mother_race + parity + ppwt + smoken, data = .x)),
    simple_model = map(train, ~ lm(bwt ~ blength + gaweeks, data = .x)),
    interaction_model = map(train, ~ lm(bwt ~ bhead * blength * baby_sex, data = .x))
  ) %>%
  mutate(
    rmse_final_model = map2_dbl(final_model, test, ~ rmse(model = .x, data = .y)),
    rmse_simple_model = map2_dbl(simple_model, test, ~ rmse(model = .x, data = .y)),
    rmse_interaction_model = map2_dbl(interaction_model, test, ~ rmse(model = .x, data = .y))
  )

# Plot RMSE distributions for each model
cv_results %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>%
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse, fill = model)) +
  geom_violin(alpha = 0.5) +
  scale_fill_manual(values = c(
    "final_model" = "#1b9e77",
    "simple_model" = "#d95f02",
    "interaction_model" = "#7570b3"
  )) +
  stat_summary(fun = "median", geom = "point", color = "black", size = 1, shape = 18) +
  labs(
    title = "RMSE Distribution by Model Type",
    x = "Model Type",
    y = "Root Mean Squared Error (RMSE)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

* Final Model (my model) has the lowest RMSE, indicating this model performs best in predicting birthweight. Simple Model (Orange) has higher RMSE compared to other models, reflecting limited predictive accuracy. Interaction Model (Purple) has slightly higher RMSE than the final model, suggesting added complexity doesn't significantly improve prediction.




